---
title: "Analysis of Weekly Newsletters: Initial Report"
output: pdf_document
---

# Executive Summary

## Questions of Interest

* 

## Statistical Analysis

* 

## Takeaways

* The next two reports, "Trends in the Open Probability" and "Trends in the Click Probability," will delve further into how various factors affect the open and click rate. 



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r}
library(ggplot2)
library(scales)
library(lubridate)
library(GGally)
library(here)
library(knitr)

i_am("initial_report/analysis_of_weekly_newsletters_initial_report.Rmd")
```

```{r}
load(here("initial_report", "weeklies1.RData"))
```

## Overview of the Data Used

Newsletter level (as opposed to subscriber level)

We focus on 54 newsletters, sent roughly weekly in the years 2019 and 2020. For this initial report, we look at aggregate statistics for the newsletters. 

<!-- We focused on the probability that a given subscriber will click on at least one link in a given newsletter. In addition to the variables used in the open probability model, -->

<!-- * Trend over the date newsletter was sent out (2019-01-01 to 2020-12-31). -->
<!-- * Whether newsletter was sent before or after start of the COVID pandemic on 2020-03-12. -->
<!-- * Time of day newsletter was sent out (6:30 am to 8:40 pm). -->
<!-- * Length of subject by number of characters. -->

<!-- we also look at the following factors related to the content within the newsletter: -->

<!-- * Number of words in newsletter. -->
<!-- * Number of links (text or image) in newsletter. -->
<!-- * Number of clickable pictures in newsletter. -->
<!-- * Number of unclickable pictures in newsletter. -->

Below are 5 example newsletters from the dataset, containing the four "content" variables mentioned above. The last variable, clicks, is the response variable of interest. 1 indicates that the subscriber clicked on any link in a given newsletter; 0 indicates that the subscriber didn't click on any link in a given newsletter.

```{r}
sample_idx <- c(5, 8, 25, 35, 47)

kable(subset(weeklies1)[sample_idx,],
      row.names = FALSE)
```



# Correlation of the Metrics

The open percentage is 

$$\mathrm{Open \ \%} = \frac{\mathrm{number \ of \ contacts \ who \ opened \ the \ email}}{\mathrm{number \ of \ contacts \ sent \ to}} \times 100\%$$

The other metrics, i.e. clicks, bounces, and unsubscribes, are similarly defined.

```{r}
metrics <- weeklies1[, names(weeklies1) %in% c("opened", "clicks", "bounces", "unsubscribes")]
```

```{r}
ggpairs(metrics, title = "Correlogram of the Metrics")
```




# Summary Statistics over Time, 2019-2020

```{r}
ggplot(data = weeklies1, aes(x = datetime, y = contacts_sent_to)) + 
  geom_point() + 
  ggtitle("Number of Contacts Sent To over Time, 2019-2020") + 
  xlab("Month/Year") +
  ylab("Number of Contacts Sent To") +
  geom_smooth() +
  scale_x_datetime(labels = date_format("%m/%y"), 
               breaks = date_breaks("2 month"))
  
```

When did the pandemic start changing things? The March 12 weekly newsletter was the first one to mention the COVID-19 pandemic and remote volunteering opportunities. 

```{r}
ggplot(data = weeklies1, aes(x = datetime, y = opened, col = covid)) + 
  geom_point() + 
  ggtitle("Open % over Time, 2019-2020") + 
  xlab("Month/Year") +
  ylab("Open %") +
  geom_smooth() +
  scale_x_datetime(labels = date_format("%m/%y"), 
               breaks = date_breaks("2 month")) + 
  geom_vline(xintercept = as.POSIXct("2020-03-01"), col = "red")
  
```

There is a spike in the open % after March. 



```{r}
ggplot(data = weeklies1, aes(x = datetime, y = clicks, col = covid)) + 
  geom_point() + 
  ggtitle("Click % over Time, 2019-2020") + 
  xlab("Month/Year") +
  ylab("Click %") +
  geom_smooth() +
  scale_x_datetime(labels = date_format("%m/%y"), 
               breaks = date_breaks("2 month")) + 
  geom_vline(xintercept = as.POSIXct("2020-03-01"), col = "red")
```



```{r}
ggplot(data = weeklies1, aes(x = datetime, y = unsubscribes, col = covid)) + 
  geom_point() + 
  ggtitle("Unsubscribe % over Time, 2019-2020") + 
  xlab("Month/Year") +
  ylab("Unsubscribe %") +
  geom_smooth() +
  scale_x_datetime(labels = date_format("%m/%y"), 
               breaks = date_breaks("2 month")) + 
  geom_vline(xintercept = as.POSIXct("2020-03-01"), col = "red")
```



# Summary Statistics by Season

```{r}
ggplot(data = weeklies1[weeklies1$covid == "Before",], aes(season, opened)) + 
  geom_boxplot()
```

```{r}
ggplot(data = weeklies1[weeklies1$covid == "After",], aes(season, opened)) + 
  geom_boxplot()
```



# Summary Statistics throughout a Day

```{r}
# extracting the time
# this sets all the dates to 2021-01-10
daytime <- as.POSIXct(format(weeklies1$datetime, format="%H:%M"),
                    tz = "America/New_York", format = "%H:%M")
```

```{r}
weeklies_before <- weeklies1[weeklies1$covid == "Before",]
weeklies_after <- weeklies1[weeklies1$covid == "After",]
```

<!-- ```{r} -->
<!-- ggplot(data = weeklies_after, aes(x = mins_since_midnight, y = opened)) +  -->
<!--   geom_point() +  -->
<!--   ggtitle("Open % throughout a Day") +  -->
<!--   xlab("Time") + -->
<!--   ylab("Open %") + -->
<!--   geom_smooth() #+ -->
<!--   #scale_x_datetime(labels = date_format("%m/%y"),  -->
<!--                #breaks = date_breaks("2 month")) -->
<!-- ``` -->

```{r}
ggplot(data = weeklies1, aes(x = daytime, y = opened, col = covid)) + 
  geom_point() + 
  ggtitle("Open % throughout a Day") + 
  xlab("Time") +
  ylab("Open %") +
  geom_smooth(aes(group = 1)) # +
  # scale_x_datetime(breaks = date_breaks("1 hour"), labels = date_format("%H:%M"))
```

There may be a slight upward trend. 



# Effect of Subject Headings

Summary Statistics for number of characters within subject heading

```{r}
summary(weeklies1$subject_length)
```

```{r}
ggplot(data = weeklies1, aes(x = subject_length, y = opened, col = covid)) + 
  geom_point() + 
  ggtitle("Open % by Subject Length") + 
  xlab("Subject Length in characters") +
  ylab("Open %") +
  geom_smooth(aes(group = 1))
```

<!-- ```{r} -->
<!-- ggplot(data = weeklies1, aes(x = subject_num_words, y = opened)) +  -->
<!--   geom_point() +  -->
<!--   ggtitle("Open % by Number of Words in Subject") +  -->
<!--   xlab("Number of Words in Subject") + -->
<!--   ylab("Open %") -->
<!-- ``` -->

Plotting by the number of words yields a similar plot. 

Small speculation: if subject heading is too long, it may not fit in the email browser and thus lead to less opens. 



<!-- # Modeling -->

<!-- ```{r} -->
<!-- weeklies1$covid <- factor(weeklies1$covid, levels = c("Before", "After")) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- opens_lm <- lm(opened ~ datetime*covid + mins_since_midnight + subject_length, data = weeklies1) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- summary(opens_lm) -->
<!-- ``` -->



<!-- Repeat for monthly donor newsletters -->




